Supporting Information for ``Nonparametric Block Bootstrap Kolmogorov-Smirnov Goodness-of-Fit Test''

### 2024-06-14 manuscript/manuscript.tex
set up proof envion in supp

\author[1]{Elizabeth Schifano}

@@

In summary, the procedure of the NPB KS test for an independent sample is
summarized as follows. Repeat the following steps for $b \in \{1, ..., B\}$.
\begin{enumerate}
\item
  Generate $X^{(b)}_1,...,X^{(b)}_n$ by sampling $X_1, \ldots, X_n$
  with replacement.
\item
  Obtain parametrically
  fitted parameters 
	$\hat\theta^{(b)}_n$ of $\theta$ from $X^{(b)}_1,...,X^{(b)}_n$.  
	%\eds{$F_\theta$ has not yet been defined.}\mc{addressed}
\item
  Obtain the empirical distribution function $F^{(b)}_n$ of
  $X^{(b)}_1,...,X^{(b)}_n$. 
\item
  Calculate bootstrap KS statistic
  \[
    T^{(b)}_n = \sup_x \Big\vert \sqrt{n}\left(F^{(b)}_n(x) 
    - F(x; \hat\theta^{(b)}_n)\right) - B_n(x) \Big\vert.
  \]
  where 
  $B_{n}(x) = \sqrt{n}(F_n(x) - F(x; \hat\theta_n))$ is the known
  bias term.
\end{enumerate}

@@

In summary, the procedure of the NPBB test is 
summarized as follows. Repeat the following steps for $b \in \{1, ..., B\}$.
\begin{enumerate}
\item
  Generate $X^{(b)}_1,...,X^{(b)}_n$ by applying moving block bootstrap 
  on the original sample as
  defined previously.
\item
  %Fit $F_\theta$ to $X^{(b)}_1,...,X^{(b)}_n$ and obtain estimate 
	%$\hat\theta^{(b)}_n$ of $\theta$.
	Obtain parametrically fitted parameters 
	$\hat\theta^{(b)}_n$ of $\theta$ from $X^{(b)}_1,...,X^{(b)}_n$.  
	\eds{changed this to remove $F_\theta$, please check}\mc{looks good}
\item
  Obtain the empirical distribution function $F^{(b)}_n$ of
  $X^{(b)}_1,...,X^{(b)}_n$. 
\item
  Calculate bootstrap KS statistic
  \[
    T^{(b)}_n = \sup_x \Big\vert \sqrt{n}\left(F^{(b)}_n(x) 
    - F(x; \hat\theta^{(b)}_n)\right) - B_n(x) \Big\vert.
  \]
  where 
  $B_{n}(x) = \sqrt{n}(E[F^{(b)}_n(x)] - 
  F(x; E[\hat\theta^{(b)}_n]))$ is the known
  bias term.
\end{enumerate}
The p-value of the block bootstrap KS test can be approximated
as $p = \sum_{b=1}^B I\{T^{(b)}_n > T_n\} / B$.

\jy{Cite Kunsch (1989)}

\jy{In the supplement, use the proof environment to give proof
  sketches; make the title ``Supplement to `paper title` by
  authors''. No need to use their template for supp.}

Assume that the time series $\{X_t\}$ is stationary and
strongly mixing. The following results provides theoretical validity
of the NPBB procedure of the KS test. Sketches of their proofs are
in the Supplementary Information, with most arguments similar to those
in \citet{kunsch1989jackknife}.

@@

These distributions
were chosen to compare results on normal and non-normal
error structures. The specific parameters were chosen because the distributions 
are very similar and their
first two moments are the same. 
For the purposes of 
evaluating if the test holds it size, when $X_t \sim N(8, 8)$, we tested that the 
marginal distribution is from the
Normal family, or 
$X_t \sim N(\cdot ; \mu, \sigma^2), \quad t = 1, \ldots, n$
for some unspecified $\mu$ and $\sigma$. When $X_t \sim \Gamma(8, 1)$, we tested
that the marginal distribution family is from the Gamma family,
or
$X_t \sim \Gamma(\cdot ; \alpha, \beta), \quad t = 1, \ldots, n$.
for some unspecified $\alpha$ and $\beta$.
%\eds{Vertical bar vs semicolon to separate data and parameter arguments.}
%\mc{changed everything to semicolon}
For the block
bootstrap step,
we used $B = 1000$ and $l = \lceil n^{1/3} \rceil$.
For each setting for $F$, $\tau$, and $n$, we replicate the method 10000 times 
to get the distribution of the p-values $p_r$ for
$r \in \{1, \dots, 10000\}$
for the test when applied to samples from the same data generating process.
%\eds{I assume $r$ is for replicate?  Please define and include how many 
%total replicates you are using.} \mc{addressed}

@@

Additional information for this article is available.
A proof of our two propositions in \ref{sub:npbb} is provided.

>>>>

\author[1]{Elizabeth D. Schifano}

@@

In summary, the procedure of the NPB KS test for an independent sample is
summarized as follows. Repeat the following steps for $b \in \{1, ..., B\}$.
\begin{enumerate}
\item
  Generate $X^{(b)}_1,...,X^{(b)}_n$ by sampling $X_1, \ldots, X_n$
  with replacement.
\item
  Obtain parametrically
  fitted parameters 
	$\hat\theta^{(b)}_n$ of $\theta$ from $X^{(b)}_1,...,X^{(b)}_n$.  
\item
  Obtain the empirical distribution function $F^{(b)}_n$ of
  $X^{(b)}_1,...,X^{(b)}_n$. 
\item
  Calculate bootstrap KS statistic
  \[
    T^{(b)}_n = \sup_x \Big\vert \sqrt{n}\left(F^{(b)}_n(x) 
    - F(x; \hat\theta^{(b)}_n)\right) - B_n(x) \Big\vert.
  \]
  where 
  $B_{n}(x) = \sqrt{n}(F_n(x) - F(x; \hat\theta_n))$ is the known
  bias term.
\end{enumerate}

@@

In summary, the procedure of the NPBB test is 
summarized as follows. Repeat the following steps for $b \in \{1, ..., B\}$.
\begin{enumerate}
\item
  Generate $X^{(b)}_1,...,X^{(b)}_n$ by applying moving block bootstrap 
  on the original sample as
  defined previously.
\item
  Obtain parametrically fitted parameters 
  $\hat\theta^{(b)}_n$ of $\theta$ from $X^{(b)}_1,...,X^{(b)}_n$.  
\item
  Obtain the empirical distribution function $F^{(b)}_n$ of
  $X^{(b)}_1,...,X^{(b)}_n$. 
\item
  Calculate bootstrap KS statistic
  \[
    T^{(b)}_n = \sup_x \Big\vert \sqrt{n}\left(F^{(b)}_n(x) 
    - F(x; \hat\theta^{(b)}_n)\right) - B_n(x) \Big\vert.
  \]
  where 
  $B_{n}(x) = \sqrt{n}(E[F^{(b)}_n(x)] - 
  F(x; E[\hat\theta^{(b)}_n]))$ is the known
  bias term.
\end{enumerate}
The p-value of the block bootstrap KS test can be approximated
as $p = \sum_{b=1}^B I\{T^{(b)}_n > T_n\} / B$.

\jy{Recover the definition of jy and eds for comments. Prepare two
  definitions in the source so that one can be easily commented out.}

\jy{Cite Kunsch (1989) in the proof as appropriate}

\jy{In the Supp: make sure notations are consistent (e.g., use
  $\{X_t: t = 1, \ldots, n\}$ for observed data);
  clean up the styles (linewidth etc.)}

Assume that the time series $\{X_t\}$ is stationary and
strongly mixing. The following results provides theoretical validity
of the NPBB procedure of the KS test. Sketches of their proofs are
in the Supporting Information, with most arguments similar to those
in \citet{kunsch1989jackknife}.

@@

These distributions
were chosen to compare results on normal and non-normal
error structures. The specific parameters were chosen because the distributions 
are very similar and their
first two moments are the same. 
For the purposes of 
evaluating if the test holds it size, when $X_t \sim N(8, 8)$, we tested that the 
marginal distribution is from the
Normal family, or 
$X_t \sim N(\cdot ; \mu, \sigma^2), \quad t = 1, \ldots, n$
for some unspecified $\mu$ and $\sigma$. When $X_t \sim \Gamma(8, 1)$, we tested
that the marginal distribution family is from the Gamma family,
or
$X_t \sim \Gamma(\cdot ; \alpha, \beta), \quad t = 1, \ldots, n$.
for some unspecified $\alpha$ and $\beta$. For the block bootstrap step,
we used $B = 1000$ and $l = \lceil n^{1/3} \rceil$.
For each setting for $F$, $\tau$, and $n$, we replicate the method 10000 times 
to get the distribution of the p-values $p_r$ for
$r \in \{1, \dots, 10000\}$
for the test when applied to samples from the same data generating process.

@@

The proofs of the two propositions in \ref{sub:npbb} are sketched in
the supplement.

### 2024-06-14 manuscript/supplement.tex
set up proof envion in supp

\documentclass[11pt]{article}

@@

\begin{document}

\section{Asymptotic limit of the KS-statistic}
We first derive the limiting distribution of the KS statistic under the null. Suppose $\hat{\theta}_n$ admits the expansion:
\begin{align}\label{eq-expan}
\hat{\theta}_n-\theta_0 = \frac{1}{n}\sum^{n}_{i=1}h(X_i;\theta_0) + \frac{1}{\sqrt{n}}R_n,    
\end{align}
where $h(\cdot;\theta_0)$ is the influence function with $\mathbb{E}[h(X_i;\theta_0)]=0$ and $R_n$ is the remainder term satisfying that
$R_n\overset{p}{\rightarrow} 0.$ Write $G_n(x):=\sqrt{n}(F_n(x)-F(x))$ and let $f(x;\theta)=\partial F(x;\theta)/\partial \theta$. We assume $\theta \in \mathbb{R}$ for ease of presentation.

@@

\section{Bootstrap consistency}
\subsection{Preliminaries}
Denote by $\mathbb{E}^*$, $\text{Var}^*$ and $\text{Cov}^*$
the expectation, variance and covariance conditional on $X_{1:n}$ in the bootstrap world. Assuming that $n=kl$ with $l$ being the block size and $k$ being the number of blocks, the bootstrap marginal distribution is 
\begin{align*}
\rho_n^*=\frac{1}{n}\sum^{k}_{j=1}\sum^{S_j+l-1}_{i=S_j} \delta_{X_i},    
\end{align*}
where $\delta_x$ denotes a point mass at $x$, and $S_1,\dots,S_k$ are i.i.d uniform on $\{1,\dots,n-l+1\}$. 
Define $G_n^*(x)$ in the same way as $G_n(x)$ with $X_{1:n}$ being replaced by the bootstrap samples $X_{1:n}^*=\{X_1^*,\dots,X_n^*\}$.

\subsection{A sketch of the proof}

We want to show that
\begin{align}\label{eq-con-boot}
\begin{pmatrix}
G_n^*(x)  \\ 
\sqrt{n}(F(x;\hat{\theta}_n^*)-F(x;\hat{\theta}_0))
\end{pmatrix}\Rightarrow^{*}
\begin{pmatrix}
G_{\infty}(x) \\
f(x;\theta_0)U
\end{pmatrix}
\end{align}
where $\Rightarrow^*$ means convergence in distribution almost surely and $\hat{\theta}_0:=\mathbb{E}^*[\hat{\theta}_n^*]$. 
In the bootstrap world, we have an expansion similar to (\ref{eq-expan}):
\begin{align*}
\hat{\theta}_n^*-\hat{\theta}_0 = \frac{1}{n}\sum^{n}_{i=1}h(X_i^*;\hat{\theta}_0) + \frac{1}{\sqrt{n}}R_n^*,    
\end{align*}
where $R_n^* \overset{p^*}{\rightarrow} 0$, where $\overset{p^*}{\rightarrow}$ means convergence in probability almost surely. Note that 
\begin{align*}
\mathbb{E}^*\left[\frac{1}{n}\sum^{n}_{i=1}h(X_i^*;\hat{\theta}_0)\right]
=&\frac{1}{n}\sum^{k}_{j=1}\mathbb{E}^*\left[\sum^{S_j+l-1}_{i=S_j}h(X_{i};\hat{\theta}_0)\right]
\\=&\frac{k}{n^2}[\sum^{n - l + 1}_{j=1}\sum^{j+l-1}_{i=j}h(X_{i};\hat{\theta}_0) +
\sum^{n}_{j=n - l + 2}[\sum^{n}_{i=j}h(X_{i};\hat{\theta}_0) +
\sum^{j - n + l -1}_{i=1}h(X_{i};\hat{\theta}_0)]]
\\=&\frac{k}{n^2}\sum^{n}_{i=1}l h(X_{i};\hat{\theta}_0),
\end{align*}.
Moreover, we have
\begin{align*}
\text{var}^*\left(\frac{1}{\sqrt{n}}\sum^{n}_{i=1}h(X_i^*;\theta_0^*)\right)
=&\text{var}^*\left(\frac{1}{\sqrt{n}}\sum^{k}_{j=1}\sum^{S_j+l-1}_{i=S_j}h(X_{i};\hat{\theta}_0)\right)
\\=&\frac{1}{l}\text{var}^*\left(\sum^{S_1+l-1}_{i=S_1}h(X_{i};\hat{\theta}_0)\right)
\\=&\frac{1}{ln}\sum^{n}_{t=1}(W_t-\bar{W})^2
\end{align*}
with $W_t=
 \begin{cases}
\sum^{t+l-1}_{i=t}h(X_i;\hat{\theta}_0)  & \text{if } t \leq n - 1 + 1,\\
\sum^{n}_{i=t}h(X_{i};\hat{\theta}_0) +
\sum^{t - n + l -1}_{i=1}h(X_{i};\hat{\theta}_0)  & \text{if } t \geq n - 1 + 2
\end{cases}$
and $\bar{W}=\sum^{n}_{t=1}W_t/n$,
which is expected to converge to $s^2$ almost surely. Similar calculations apply to $G_n^*(x)$ and we expect that
\begin{align*}
&\text{Cov}^*(G_n^*(x),G_n^*(x'))\overset{a.s.}{\rightarrow}   \sum^{+\infty}_{h=-\infty}\text{Cov}(\mathbf{1}\{X_{h}\leq x\},\mathbf{1}\{X_0\leq x'\}),\\
&\text{Cov}^*\left(\frac{1}{\sqrt{n}}\sum^{n}_{i=1}h(X_i^*;\hat{\theta}_0), G_n^*(x)\right)\overset{a.s.}{\rightarrow} \sum^{+\infty}_{h=-\infty}\text{Cov}(h(X_{h},\theta_0),\mathbf{1}\{X_0\leq x\}).
\end{align*}
By showing the finite-dimensional convergence and tightness in the bootstrap world (conditional on the data $X_{1:n}$), we can establish \eqref{eq-con-boot}.
From the CLT, the process $G_n^*(x) = \sqrt{n}(F_n^*(x) - F_n(x))$ converges 
to 
$G_{\infty}(x)$.

@@

The term $\frac{1}{\sqrt{n}} \sum_{i=1}^{n} h(X_i^*; \hat{\theta}_0)$ 
converges in distribution to a normal random variable $U$ with variance 
$s^2$.
We can show tightness of $G_n^*(x)$ similarly to how we established 
tightness of $G_n(x)$
The results then follow from the continuous mapping theorem.

>>>>

\documentclass[12pt]{article}

@@

\title{Supporting Information for
  ``Nonparametric Block Bootstrap Kolmogorov-Smirnov Goodness-of-Fit
  Test''}
\author{Mathew Chandy, Elizabeth D. Schifano, Jun Yan, and Xianyang
  Zhang}

\begin{document}
\maketitle

\section{Asymptotic Limit of the KS Statistic}

\begin{proof}[Proof of Proposition~1]
We first derive the limiting distribution of the KS statistic under the null. Suppose $\hat{\theta}_n$ admits the expansion:
\begin{align}\label{eq-expan}
\hat{\theta}_n-\theta_0 = \frac{1}{n}\sum^{n}_{i=1}h(X_i;\theta_0) + \frac{1}{\sqrt{n}}R_n,    
\end{align}
where $h(\cdot;\theta_0)$ is the influence function with $\mathbb{E}[h(X_i;\theta_0)]=0$ and $R_n$ is the remainder term satisfying that
$R_n\overset{p}{\rightarrow} 0.$ Write $G_n(x):=\sqrt{n}(F_n(x)-F(x))$ and let $f(x;\theta)=\partial F(x;\theta)/\partial \theta$. We assume $\theta \in \mathbb{R}$ for ease of presentation.

@@

\end{proof}

\section{Bootstrap Consistency}

Denote by $\mathbb{E}^*$, $\text{Var}^*$ and $\text{Cov}^*$
the expectation, variance and covariance conditional on $X_{1:n}$ in the bootstrap world. Assuming that $n=kl$ with $l$ being the block size and $k$ being the number of blocks, the bootstrap marginal distribution is 
\begin{align*}
\rho_n^*=\frac{1}{n}\sum^{k}_{j=1}\sum^{S_j+l-1}_{i=S_j} \delta_{X_i},    
\end{align*}
where $\delta_x$ denotes a point mass at $x$, and $S_1,\dots,S_k$ are i.i.d uniform on $\{1,\dots,n-l+1\}$. 
Define $G_n^*(x)$ in the same way as $G_n(x)$ with $X_{1:n}$ being replaced by the bootstrap samples $X_{1:n}^*=\{X_1^*,\dots,X_n^*\}$.

\begin{proof}]Proof of Proposition~2]
    We want to show that
\begin{align}\label{eq-con-boot}
\begin{pmatrix}
G_n^*(x)  \\ 
\sqrt{n}(F(x;\hat{\theta}_n^*)-F(x;\hat{\theta}_0))
\end{pmatrix}\Rightarrow^{*}
\begin{pmatrix}
G_{\infty}(x) \\
f(x;\theta_0)U
\end{pmatrix}
\end{align}
where $\Rightarrow^*$ means convergence in distribution almost surely and $\hat{\theta}_0:=\mathbb{E}^*[\hat{\theta}_n^*]$. 
In the bootstrap world, we have an expansion similar to (\ref{eq-expan}):
\begin{align*}
\hat{\theta}_n^*-\hat{\theta}_0 = \frac{1}{n}\sum^{n}_{i=1}h(X_i^*;\hat{\theta}_0) + \frac{1}{\sqrt{n}}R_n^*,    
\end{align*}
where $R_n^* \overset{p^*}{\rightarrow} 0$, where $\overset{p^*}{\rightarrow}$ means convergence in probability almost surely. Note that 
\begin{align*}
\mathbb{E}^*\left[\frac{1}{n}\sum^{n}_{i=1}h(X_i^*;\hat{\theta}_0)\right]
=&\frac{1}{n}\sum^{k}_{j=1}\mathbb{E}^*\left[\sum^{S_j+l-1}_{i=S_j}h(X_{i};\hat{\theta}_0)\right]
\\=&\frac{k}{n^2}[\sum^{n - l + 1}_{j=1}\sum^{j+l-1}_{i=j}h(X_{i};\hat{\theta}_0) +
\sum^{n}_{j=n - l + 2}[\sum^{n}_{i=j}h(X_{i};\hat{\theta}_0) +
\sum^{j - n + l -1}_{i=1}h(X_{i};\hat{\theta}_0)]]
\\=&\frac{k}{n^2}\sum^{n}_{i=1}l h(X_{i};\hat{\theta}_0),
\end{align*}.
Moreover, we have
\begin{align*}
\text{var}^*\left(\frac{1}{\sqrt{n}}\sum^{n}_{i=1}h(X_i^*;\theta_0^*)\right)
=&\text{var}^*\left(\frac{1}{\sqrt{n}}\sum^{k}_{j=1}\sum^{S_j+l-1}_{i=S_j}h(X_{i};\hat{\theta}_0)\right)
\\=&\frac{1}{l}\text{var}^*\left(\sum^{S_1+l-1}_{i=S_1}h(X_{i};\hat{\theta}_0)\right)
\\=&\frac{1}{ln}\sum^{n}_{t=1}(W_t-\bar{W})^2
\end{align*}
with $W_t=
 \begin{cases}
\sum^{t+l-1}_{i=t}h(X_i;\hat{\theta}_0)  & \text{if } t \leq n - 1 + 1,\\
\sum^{n}_{i=t}h(X_{i};\hat{\theta}_0) +
\sum^{t - n + l -1}_{i=1}h(X_{i};\hat{\theta}_0)  & \text{if } t \geq n - 1 + 2
\end{cases}$
and $\bar{W}=\sum^{n}_{t=1}W_t/n$,
which is expected to converge to $s^2$ almost surely. Similar calculations apply to $G_n^*(x)$ and we expect that
\begin{align*}
&\text{Cov}^*(G_n^*(x),G_n^*(x'))\overset{a.s.}{\rightarrow}   \sum^{+\infty}_{h=-\infty}\text{Cov}(\mathbf{1}\{X_{h}\leq x\},\mathbf{1}\{X_0\leq x'\}),\\
&\text{Cov}^*\left(\frac{1}{\sqrt{n}}\sum^{n}_{i=1}h(X_i^*;\hat{\theta}_0), G_n^*(x)\right)\overset{a.s.}{\rightarrow} \sum^{+\infty}_{h=-\infty}\text{Cov}(h(X_{h},\theta_0),\mathbf{1}\{X_0\leq x\}).
\end{align*}
By showing the finite-dimensional convergence and tightness in the bootstrap world (conditional on the data $X_{1:n}$), we can establish \eqref{eq-con-boot}.
From the CLT, the process $G_n^*(x) = \sqrt{n}(F_n^*(x) - F_n(x))$ converges 
to 
$G_{\infty}(x)$.

@@

The term $\frac{1}{\sqrt{n}} \sum_{i=1}^{n} h(X_i^*; \hat{\theta}_0)$ 
converges in distribution to a normal random variable $U$ with variance 
$s^2$.
We can show tightness of $G_n^*(x)$ similarly to how we established 
tightness of $G_n(x)$
The results then follow from the continuous mapping theorem.
\end{proof}
